The assessment of the ability of measurements, nowcasts, and forecasts to track changes

### 2024-04-15 paper.tex
Work on BCS todos

%%% --- Commenting, Todonotes
\usepackage[colorinlistoftodos,prependcaption]{todonotes}
%\usepackage{comment, soul}
%\newcommand{\ar}{$\rightarrow$}
\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}

@@

%%% --- Acronyms
\usepackage{acro}
\DeclareAcronym{kde}{short=KDE, long=kernel density estimation}
\DeclareAcronym{bca}{short=BCa, long=bias-corrected and accelerated}
\DeclareAcronym{cdf}{short=CDF, long=cumulative distribution function}
\DeclareAcronym{pdf}{short=PDF, long=probability density function}
\DeclareAcronym{nbp}{short=NBP, long=non-invasive blood pressure}
\DeclareAcronym{abp}{short=ABP, long=invasive arterial blood pressure}
\DeclareAcronym{bs}{short=BS, long=Brier score}
\DeclareAcronym{rmse}{short=RMSE, long=root mean square error}
\DeclareAcronym{mae}{short=MAE, long=mean absolute error}
\DeclareAcronym{tadda}{short=TADDA, long=targeted absolute deviation with direction augmentation}
\DeclareAcronym{atc}{short=ATC, long=ability to track changes}
\DeclareAcronym{mnf}{short=MNF, long={measurement, nowcast, or forecast}}

@@

\section{Introduction}\label{sec:atc-introduction}
BCS: We need to avoid non-standard abbreviations that confuse readers such as MNF. Please delete the abbreviation MNF throughout the text.

Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest. It is thus important to meticulously assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the ability to track changes (\ac{atc}). Although measurements, nowcasts, and forecasts fundamentally differ as they either measure or predict a value, similar methods can be used to assess their \ac{atc}.

@@

Measurements aim to obtain accurate and precise values of a measureable quantity (measureand).
BCS: Please cite https://pubmed.ncbi.nlm.nih.gov/32185615/ and https://pubmed.ncbi.nlm.nih.gov/32185616/
Repeated measurements can be used to track changes in a value over time.  
When introducing new measurement methods, they are evaluated against current reference methods, often called the "gold standard", by simultaneously measuring the same quantity with the new method and the reference method – often in various individuals or different clinical settings.

@@

In a prediction competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error [BCS: Please cite: https://pubmed.ncbi.nlm.nih.gov/12578081/] are commonly used.
How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.

@@

The assessment of \ac{atc} is based on the measured and predicted changes in a value of interest over a time horizon $l$.
The \textit{measured} change is straightforward to compute for all types of \ac{mnf}.
Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
\begin{equation}\label{eq:diffy}
    \diffylt = (y_{t} - y_{t-l}) \quad \text{for}\ t = l, \dots, T.
\end{equation}

@@

The definition of the \textit{predicted} change depends on the context;
Table~\ref{tab:notation} summarizes the notation for \ac{mnf} and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
We call $t$ the \textit{target time} and $\tau$ the \textit{issue time}.
The predicted change is computed by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet.
If the true value $y_{t-l}$ is known at time $t$, the predicted change is computed by the difference between the nowcast and the true value, as $y_{t-l}$ is also known by the nowcaster and incorporated into the nowcast.
Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.

@@

In applications, data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
We refer to time steps for which either \ac{mnf} or true values are unavailable as missing values.
Systematical missing values could lead to a biased assessment, and missing data should be inspected for any underlying patterns.
If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
Note that in the case of measurement data, one missing value in the time series leads to two undefined differences in the change series; that is if $x_t$ is missing, $\diffxt$ and $\diffxt[t+l]$ are undefined; if an observation $y_t$ is missing, $\diffyt$ and $\diffyt[t+l]$ are undefined.
The data pair is excluded even if the corresponding nowcast or forecast is available.

@@

Formally, the assessment of \ac{atc} is the same for \ac{mnf}, given the notation for the respective application of Section~\ref{subsec:notation}.
In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
The \ac{atc} is maximal if all predicted change directions are correct; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
Consequently, when assessing the \ac{atc}, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998, Saugel2015}.
In a four-quadrant plot, the occured changes and the predicted changes are plotted together, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrectly predicted changes.
Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.

@@

Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:atc-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015, Critchley2010}.
Points within that area are omitted in the calculation of the measures.
In particular, the \ac{mnf} are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
The measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} without points in the exclusion area $E$ are
\begin{align}
    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}},\label{eq:acceps}\\
    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}. \label{eq:accmeps}
\end{align}
The measures are then estimators for the probability of predicting the correct direction, given that the point's location is not driven by noise or non-informative changes.

The estimators accept various shapes of the exclusion area (see Figure~\ref{fig:aatc_4q}).
A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
%Points are included if the true value or prediction is unlikely to be noise-driven.
%In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
%At least for \ac{mnf}, small amounts of noise are inevitable.
A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
%A point is only included if no component is likely to be driven by noise.
%For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.

@@

The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the \ac{mnf}.
Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
Thus, if a change of $\xcond$ is observed in practice, one can directly assess its credibility.
A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
\begin{align*}
P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond) = \begin{cases}
                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond < 0, \\
                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond > 0, \\
\end{cases}
\end{align*}
for $\xcond \neq 0$ through a \ac{kde}.
\textcite{Gramacki2018} provides a comprehensive introduction to multivariate \ac{kde}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ for all values of $\xcond \in \R$.
Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters.
We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwidth selector (see Appendix~\ref{subsec:appendix-kde}).

@@

We here consider the \ac{atc} of non-invasive blood pressure measurements from the MIMIC-III database that includes of critically ill patients treated in intensive care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
We focus on \acf{abp} and \acf{nbp} measurements and thus limit our analysis to datasets containing simultaneous measurements of \ac{abp} and \ac{nbp} simultaneously. 2,548 (BCS: what? datasets? patients?) include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
We assess the \ac{atc} of non-invasive blood pressure measurements (test emthod) compared to intraarterial blood pressure measurements (referecne method, gold standard).
We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.

@@

In this paper, we examine various methods to assess the \acf{atc} for \acf{mnf}, that is, whether it predicts the correct direction of changes.
BCS: Hier einfach wieder gleiches Wording wie in Einleitung etc. Sonst verwirren wir die Leser
While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
The \ac{atc} assessment can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.

Four-quadrant plots facilitate the visual inspection of the \ac{atc} for a \ac{mnf} (see Section~\ref{subsec:aatc-four-quadrant-plot}).
The \ac{atc} ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates \ac{atc}.
Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:atc-measures}).
The positive and negative \ac{atc} ratios analyze the \ac{atc} ratio given whether the predicted change is positive or negative, respectively.
Thus, they quantify the credibility of the respective predictions.
The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative \ac{atc} and that they add valuable information to the \ac{atc} ratio.
In the applications, the bootstrap confidence intervals of Section~\ref{subsec:atc-measures} are used to quantify the estimation uncertainty of the \ac{atc} measures.
The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
For models with reasonably high \ac{atc}, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high \ac{atc} differences.

A conditional \ac{atc} plot visualizes the probability of correct change direction prediction over the predicted change of the \ac{mnf} (see Section~\ref{subsec:aatc-cond-prob}).
It is based on a multivariate \acf{kde} of predicted and observed change.
In the application, the conditional \ac{atc} plot gives reasonable insights into the local effects of the \ac{atc}.
Section~\ref{subsec:aatc-probabilistic} adapts measures of probabilistic forecast evaluation to the \ac{atc} assessment of probabilistic forecasts and nowcasts.
The \acf{bs} as numerical assessment of probabilistic \ac{atc} is introduced, and reliability diagrams are used to visualize the local \ac{atc} of probabilistic forecasts.

>>>>

%%% --- Commenting, Todonotes
%\usepackage[colorinlistoftodos,prependcaption]{todonotes}
%%\usepackage{comment, soul}
%%\newcommand{\ar}{$\rightarrow$}
%\presetkeys{todonotes}{backgroundcolor=yellow!30, bordercolor=yellow!50, linecolor=yellow!50, figwidth=\textwidth}{}
%\newcommand{\hl}[1]{\textcolor{Aquamarine}{#1}}

@@

%%% --- Acronyms
\usepackage{acro}
\DeclareAcronym{kde}{short=KDE, long=kernel density estimation}
\DeclareAcronym{bca}{short=BCa, long=bias-corrected and accelerated}
\DeclareAcronym{cdf}{short=CDF, long=cumulative distribution function}
\DeclareAcronym{pdf}{short=PDF, long=probability density function}
\DeclareAcronym{nbp}{short=NBP, long=non-invasive blood pressure}
\DeclareAcronym{abp}{short=ABP, long=invasive arterial blood pressure}
\DeclareAcronym{bs}{short=BS, long=Brier score}
\DeclareAcronym{rmse}{short=RMSE, long=root mean square error}
\DeclareAcronym{mae}{short=MAE, long=mean absolute error}
\DeclareAcronym{tadda}{short=TADDA, long=targeted absolute deviation with direction augmentation}
\DeclareAcronym{atc}{short=ATC, long=ability to track changes}

@@

\section{Introduction}\label{sec:atc-introduction}

Measurements, nowcasts, or forecasts ideally should correctly reflect changes in the values of interest. It is thus important to meticulously assess the ability of measurements, nowcasts, or forecasts to correctly predict the direction of changes in values – which we refer to as the \acf{atc}. Although measurements, nowcasts, and forecasts fundamentally differ as they either measure or predict a value, similar methods can be used to assess their \ac{atc}.

@@

Measurements aim to obtain accurate and precise values of a measureable quantity (measureand; \citealp{Squara2021,Squara2021a}).
Repeated measurements can be used to track changes in a value over time.
When introducing new measurement methods, they are evaluated against current reference methods, often called the \enquote{gold standard}, by simultaneously measuring the same quantity with the new method and the reference method – often in various individuals or different clinical settings.

@@

In a prediction competition on armed conflicts, the assessment of the \ac{atc} recently gained attention as \citet{Vesco2022} proposed the novel \ac{tadda} score with an additive tracking-changes-component for evaluation.
However, the score poses an unintuitive incentive to forecasters and is thus theoretically problematic~\parencite{Bracher2023}.
When evaluating the measurment performance of a measurement method, comparative statistics such as Bland-Altman analysis~\citep{Bland1986} and the percentage error \parencite{Critchley1999} are commonly used.
How to best assess the \ac{atc} of measurement methods is a field of ongoing research~\citep{Saugel2015, Saugel2018, Hiraishi2021}.

@@

The assessment of \ac{atc} is based on the measured and predicted changes in a value of interest over a time horizon $l$.
The \textit{measured} change is straightforward to compute for all types of measurement, nowcast, or forecast.
Let $\mathbf{y} = (y_t)_{t=0}^T$ denote the actual values for nowcasting or forecasting, or gold standard measurements up to time $T$.
The sequence of observed changes is then given by the differences of values in $\mathbf{y}$ with horizon $l$, that is,
\begin{equation}\label{eq:diffy}
    \diffylt = (y_{t} - y_{t-l}) \quad \text{for}\ t = l, \dots, T.
\end{equation}

@@

The definition of the \textit{predicted} change depends on the context;
Table~\ref{tab:notation} summarizes the notation for measurements, nowcasts, or forecasts and the computation of the predicted change.
For nowcasting, let $x_{t | \tau}$ denote the nowcast for time $t$ computed with the knowledge of time $\tau$.
We call $t$ the \textit{target time} and $\tau$ the \textit{issue time}.
The predicted change is computed by
\begin{equation}
  \diffxl =
\begin{cases}
(x_{t|t} - x_{t-l|t})^T_{t=l} & \text{if } y_{t-l} \text{ is not known at time } t, \\
(x_{t|t} - y_{t-l})^T_{t=l} & \text{otherwise}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}
When computing the predicted change of a nowcast for a time $t$, we use the best knowledge available at that time $t$, and the true value might not be known yet.
If the true value $y_{t-l}$ is known at time $t$, the predicted change is computed by the difference between the nowcast and the true value, as $y_{t-l}$ is also known by the nowcaster and incorporated into the nowcast.
Through the computation in Equation~\eqref{eq:diffxl_nowcasting}, the predicted change can be computed with the knowledge of the nowcaster at time $t$.

@@

In applications, data are often not available for all time steps, for example, due to technical problems or delays in data transfer (see the examples in Sections~\ref{sec:application-covid} and~\ref{sec:application-eda}).
We refer to time steps for which either measurement, nowcast, or forecast or true values are unavailable as missing values.
Systematical missing values could lead to a biased assessment, and missing data should be inspected for any underlying patterns.
If the missing values are not systematic, random, and occur scarcely, data pairs with missing values can be excluded from the data to calculate the measures~\parencite[see][Section 1.3]{VanBuuren2018}.
Note that in the case of measurement data, one missing value in the time series leads to two undefined differences in the change series; that is if $x_t$ is missing, $\diffxt$ and $\diffxt[t+l]$ are undefined; if an observation $y_t$ is missing, $\diffyt$ and $\diffyt[t+l]$ are undefined.
The data pair is excluded even if the corresponding nowcast or forecast is available.

@@

Formally, the assessment of \ac{atc} is the same for measurement, nowcast, or forecast, given the notation for the respective application of Section~\ref{subsec:notation}.
In the following, we omit the horizon $l$ for ease of notation; $\diffx$ and $\diffy$ refer to $\diffxl$ and $\diffyl$ for a common horizon $l$.
The \ac{atc} is maximal if all predicted change directions are correct; that is, the sign of all elements of $\diffx$ and $\diffy$ coincide.
Consequently, when assessing the \ac{atc}, we examine the statistical consistency of $\sign(\diffx)$ and $\sign(\diffy)$.
A simple yet insightful method is the four-quadrant plot~\citep[see, e.g., ][]{Perrino1998, Saugel2015}.
In a four-quadrant plot, the occured changes and the predicted changes are plotted together, that is, $(\diffyt, \diffxt)$ for $t = l, \dots, T$.
Thus, the x-axis of a four-quadrant plot shows the true value differences, whereas the y-axis displays the prediction data differences.
Points in the green upper right and lower left quadrants reflect a correct change direction for the respective time step, whereas points in the remaining red quadrants show incorrectly predicted changes.
Figure~\ref{fig:aatc_basic_4q} displays a basic four-quadrant plot and Figure~\ref{fig:aatc_basic_4q_sample} shows a four-quadrant graph for simulated data with $T=1461$, for example, four years of daily data.

@@

Using an exclusion area around the zero point, as further defined below, is a straightforward and highly interpretable extension of the measures of Section~\ref{subsec:atc-measures} accounting for such effects~\citep[see, e.g.,][]{Saugel2015, Critchley2010}.
Points within that area are omitted in the calculation of the measures.
In particular, the measurement, nowcast, or forecast are likely to have a noise component; thus, $\diffx$ should be subject to an exclusion area.
The measures of Equations~\eqref{eq:acc},~\eqref{eq:accp} and~\eqref{eq:accm} without points in the exclusion area $E$ are
\begin{align}
    \acceps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffx \diffy > 0} \ind{(\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{(\diffyt, \diffxt) \notin E}},\label{eq:acceps}\\
    \accpeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt > 0, (\diffyt, \diffxt) \notin E}}, \ \text{and} \label{eq:accpeps}\\
    \accmeps (\diffx, \diffy, E) &\coloneqq \frac{\sum_{t \in \mathcal{T}} \ind{\diffxt \diffyt > 0} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}{\sum_{t \in \mathcal{T}} \ind{\diffxt < 0, (\diffyt, \diffxt) \notin E}}. \label{eq:accmeps}
\end{align}
The measures are then estimators for the probability of predicting the correct direction, given that the point's location is not driven by noise or non-informative changes.

The estimators accept various shapes of the exclusion area (see Figure~\ref{fig:aatc_4q}).
A rectangular exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \land (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, leaves out points that are small in both components.
%Points are included if the true value or prediction is unlikely to be noise-driven.
%In the example graph in Figure~\ref{fig:aatc_basic_4q_excl_box}, only point 1 is excluded.
An exclusion area along one axis, for example, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x)\}$ for $\varepsilon_x > 0$, removes points in which one of the components could change sign by a small amount of noise.
%At least for measurement, nowcast, or forecast, small amounts of noise are inevitable.
A cross-shaped exclusion area, $E = \{(y, x) \in \R^2: (-\varepsilon_x \leq x \leq \varepsilon_x) \lor (-\varepsilon_y \leq y \leq \varepsilon_y) \}$ for $\varepsilon_x, \varepsilon_y > 0$, along both axes accounts for the sign reversal in both components.
%A point is only included if no component is likely to be driven by noise.
%For the latter two methods, points 1 and 7 in Figure~\ref{fig:aatc_basic_4q_excl_axis} or 1, 6, and 7 are excluded in Figure~\ref{fig:aatc_basic_4q_excl_cross}, respectively.

@@

The estimators described above give information on the probabilities $P(\diffxrv \diffyrv > 0 | \diffxrv \diffyrv \notin E)$, $P(\diffxrv \diffyrv > 0 | \diffxrv > 0, \diffxrv \diffyrv \notin E)$ and $P(\diffxrv \diffyrv > 0 | \diffxrv < 0, \diffxrv \diffyrv \notin E)$.
A still finer analysis might be gained by considering the conditional distribution $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ to assess the \ac{atc} of a prediction for a specific change $\diffxrv = \xcond$ of the measurement, nowcast, or forecast.
Thereby, $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ denotes the probability of a correct direction given a predicted change of $\xcond$.
Thus, if a change of $\xcond$ is observed in practice, one can directly assess its credibility.
A multivariate \acf{kde} facilitates a continuous estimation of $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ by estimating the components $f_{\diffxrv, \diffyrv}$ and $f_{\diffxrv}$ of
\begin{align*}
P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond) = \begin{cases}
                                              \int_{-\infty}^0 \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond < 0, \\
                                              \int_{0}^{\infty} \frac{f_{\diffxrv, \diffyrv}(\xcond, y)}{f_{\diffxrv}(\xcond)} \ \textrm{d} \: y & \text{if } \xcond > 0, \\
\end{cases}
\end{align*}
for $\xcond \neq 0$ through a \ac{kde}.
\textcite{Gramacki2018} provides a comprehensive introduction to multivariate \ac{kde}, and implementations are available in many programming languages, for example, in the \verb|statsmodels| in Python~\parencite{Seabold2010}.
The \ac{kde} yields estimates for $P(\diffxrv \diffyrv > 0 | \diffxrv = \xcond)$ for all values of $\xcond \in \R$.
Multivariate \ac{kde} takes a kernel and bandwidth selector as modeling parameters.
We advise using a Gaussian kernel and the cross-validation maximum likelihood as bandwidth selector (see Appendix~\ref{subsec:appendix-kde}).

@@

We here consider the \ac{atc} of non-invasive blood pressure measurements from the MIMIC-III database that includes of critically ill patients treated in intensive care units of the Beth Israel Deaconess Medical Center in Boston (Massachusetts, USA, \cite{Johnson2016}, \citealp{Moody2017}; available through \citealp{Goldberger2000}).
We focus on \acf{abp} and \acf{nbp} measurements and thus limit our analysis to datasets containing simultaneous measurements of \ac{abp} and \ac{nbp} simultaneously. 2,548 datasets include at least one measurement of systolic \ac{abp} and \ac{nbp} and 1,327 include at least one pair of simultaneously measured systolic \ac{abp} and \ac{nbp}; for the mean \ac{abp} and \ac{nbp}, the numbers are 2,605 and 1,516, respectively.
We assess the \ac{atc} of non-invasive blood pressure measurements (test emthod) compared to intraarterial blood pressure measurements (referecne method, gold standard).
We consider the horizons of one minute, five minutes, and 15 minutes for the \ac{atc} assessment, as those are typical intervals of NBP measurements.

@@

In this paper, we examine various methods to assess the \acf{atc} for measurements, nowcasts, or forecasts, that is, whether they correctly predict the direction of changes in values.
While the computation of predicted change varies between the application areas measurement, nowcasting, and forecasting, the predict the correct direction of changes assessment using the computed predicted and observed change is similar.
The \ac{atc} assessment can accompany other evaluation techniques, such as measures of deviation or probabilistic scoring rules.

Four-quadrant plots facilitate the visual inspection of the \ac{atc} for a measurement, nowcast, or forecast (see Section~\ref{subsec:aatc-four-quadrant-plot}).
The \ac{atc} ratio, the ratio of change directions predicted correctly over the total number of changes, numerically evaluates \ac{atc}.
Visually, it is the proportion of concordant points in a four-quadrant plot (see Section~\ref{subsec:atc-measures}).
The positive and negative \ac{atc} ratios analyze the \ac{atc} ratio given whether the predicted change is positive or negative, respectively.
Thus, they quantify the credibility of the respective predictions.
The applications of Section~\ref{sec:application_measurement} show that models, in general, indeed have different positive and negative \ac{atc} and that they add valuable information to the \ac{atc} ratio.
In the applications, the bootstrap confidence intervals of Section~\ref{subsec:atc-measures} are used to quantify the estimation uncertainty of the \ac{atc} measures.
The width of the confidence intervals is around 0.1 for around 100 samples, while it is around 0.01 for 8000 samples.
For models with reasonably high \ac{atc}, 100 samples are thus sufficient to differentiate from random guessing or to assess models with high \ac{atc} differences.

A conditional \ac{atc} plot visualizes the probability of correct change direction prediction over the predicted change of the measurement, nowcast, or forecast (see Section~\ref{subsec:aatc-cond-prob}).
It is based on a multivariate \acf{kde} of predicted and observed change.
In the application, the conditional \ac{atc} plot gives reasonable insights into the local effects of the \ac{atc}.
Section~\ref{subsec:aatc-probabilistic} adapts measures of probabilistic forecast evaluation to the \ac{atc} assessment of probabilistic forecasts and nowcasts.
The \acf{bs} as numerical assessment of probabilistic \ac{atc} is introduced, and reliability diagrams are used to visualize the local \ac{atc} of probabilistic forecasts.
